Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2022-09-12 manuscript.Rmd
move plot to own script

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
output:
  bookdown::pdf_document2:
    latex_engine: "xelatex"
    toc: false
    number_sections: false
header-includes:
   - \usepackage{libertine}
fontsize: 11pt
linestretch: 1.2
geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"
papersize: a4
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
csl: apa.csl
repro:
  files: abstract.md
  packages:
    - tidyverse
    - ggplot2
    - ggthemes
    - patchwork
    - tikzDevice
  tlmgr:
    - pgf
    - preview
    - libertine
  apt:
    - git
    - rsync
    - curl
    - inkscape
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE
)

@@

```{r, include=FALSE}
bayes <- function(h, eh, enh)(h * eh)/((h * eh) + ((1 - h) * enh))
```

@@

```{r, include=FALSE}
forbid_additional_args <- function(...)stopifnot(length(list(...)) == 0L)

#' Apply the Bayes Theorem
#'
#' @param h The prior probability of the hypothesis \eqn{P(H)}.
#' @param detectability The probability of detecting evidence, if hypothesis holds \eqn{P(E|H)}.
#' @param theoretical_risk The probability of not detecting evidence, if hypothesis does not hold \eqn{P(¬E|¬H)}.
#' @param tau Same as `theoretical_risk` for compatibility.
#' @return The posterior probability of the hypothesis after observing evidence \eqn{P(H|E)}.
bayes <- function(h, detectability, theoretical_risk, ...) {
  forbid_additional_args(...)
  true_positive <- h * detectability
  false_positive <- (1 - h) * (1 - theoretical_risk)
  true_positive / (true_positive + false_positive)
}

# all formulas taken from Bayesian Philosophy of Science, p. 51
# first letter = function name in Bayesian Philosophy of Science
# optional letter = p for prime
# second letter = m for measure
measures <- list(
  pm = list(
    name = "Posterior Probability",
    formula = "\\frac{P(H)P(E|H)}{P(H)P(E|H) + P(¬H)P(E|¬H)}",
    fun = bayes
  ),
  dm = list(
    name = "Difference Measure",
    formula = "d(H, E) = p(H|E) - p(H)",
    fun = function(h, detectability, theoretical_risk, ...) {
      forbid_additional_args(...)
      he <- bayes(h, detectability, theoretical_risk)
      he - h
    }
  ),
  rm = list(
    name = "Log-Ratio Measure",
    formula = "r(H, E) = log(\\frac{p(H|E)}{p(H)})",
    fun = function(h, detectability, theoretical_risk, ...) {
      forbid_additional_args(...)
      he <- bayes(h, detectability, theoretical_risk)
      log(he / h)
    }
  ),
  lm = list(
    name = "Log-Likelihood Measure",
    formula = "l(H, E) = log(\\frac{p(H|E)}{p(E|\\neg{}H)})",
    fun = function(h, detectability, theoretical_risk, ...) {
      forbid_additional_args(...)
      he <- bayes(h, detectability, theoretical_risk)
      enh <- 1 - theoretical_risk
      log(he / enh)
    }
  ),
  km = list(
    name = "Kemeny–Oppenheim Measure",
    formula = "k(H, E) = \\frac{p(E|H) - p(E|\\neg{}H)}{p(E|H) + p(E|\\neg{}H)}",
    fun = function(h, detectability, theoretical_risk, ...) {
      forbid_additional_args(...)
      eh <- detectability
      enh <- 1 - theoretical_risk
      (eh - enh) / (eh + enh)
    }
  ),
  zm = list(
    name = "Generalized Entailment Measure",
    formula =
      "z(H, E) =\\begin{cases}
  \\frac{p(E|H) - p(H)}{1 - p(H)}, & \\text{if } p(H|E) \\geq{} p(H) \\\\
  \\frac{p(E|H) - p(H)}{p(H)}, & \\text{if } p(H|E) < p(H)
\\end{cases}",
    fun = function(h, detectability, theoretical_risk, ...) {
      forbid_additional_args(...)
      he <- bayes(h, detectability, theoretical_risk)
      ifelse(he >= h, (he - h) / (1 - h), (he - h) / h)
    }
  ),
  sm = list(
    name = "Christensen–Joyce Measure",
    formula = "s(H, E) = p(H|E) - p(H|\\neg{}E)",
    fun = function(h, detectability, theoretical_risk, ...) {
      forbid_additional_args(...)
      he <- bayes(h, detectability, theoretical_risk)
      enh <- 1 - theoretical_risk
      eh <- detectability
      e <- h * eh + (1 - h) * enh
      hne <- (h * (1 - eh)) / (1 - e)
      he - hne
    }
  ),
  cpm = list(
    name = "Carnap’s Relevance Measure",
    formula = "c'(H, E) = p(E)(p(H|E) - p(H))",
    fun = function(h, detectability, theoretical_risk, ...) {
      forbid_additional_args(...)
      he <- bayes(h, detectability, theoretical_risk)
      enh <- 1 - theoretical_risk
      e <- h * detectability + (1 - h) * enh
      e * (he - h)
    }
  ),
  rpm = list(
    name = "Rips Measure",
    formula = "r'(H, E) = 1 - \\frac{p(\\neg{}H|E)}{p(\\neg{}H)}",
    fun = function(h, detectability, theoretical_risk, ...) {
      forbid_additional_args(...)
      he <- bayes(h, detectability, theoretical_risk)
      1 - (1 - he) / (1 - h)
    }
  )
)

measures <- tibble(!!!transpose(measures)) %>% 
  mutate(name = simplify(name, "chr"),
         formula = simplify(formula, "chr"))

#' Integrate over uncertain theoretical risk
#' 
#' If theoretical risk is maximally uncertain, we represent this uncertainty
#' as a uniform distribution from zero to tau from which we sample and average the measures.
#'
#' @param tau The guaranteed upper bound of theoretical risk.
#' @param n Number of samples from the uniform distribution.
#' @param seed Optional seed to set before execution. Is reset to global state on exit.
#' @return A function similar to the original
uncertainly <- function(measure, tau, n = 1e5, seed = NULL){
  function(h, detectability, ...){
    forbid_additional_args(...)
    if(!is.null(seed)){
      if(exists(".Random.seed"))on.exit(set.seed(.Random.seed))
      set.seed(seed)
    }
    theoretical_risk_samples <- runif(n, 0, tau)
    mean(measure(h, detectability, theoretical_risk_samples))
  }
}
certainly <- function(measure, tau){
  function(h, detectability, ...){
    forbid_additional_args(...)
    measure(h, detectability, tau)
  }
}
```

```{r, cache=TRUE, echo=FALSE}
step <- 0.01

grid <-
  tidyr::expand_grid(
    h = .1,
    detectability = .8,
    tau = seq(step, 1 - step, step)
  )
apply_on_grid <- function(f, grid, seed = NULL, n = 1e4) {
  certain <- function(h, detectability, tau, ...){
    forbid_additional_args(...)
    certainly(f, tau)(h, detectability)
  }
  uncertain <- function(h, detectability, tau, ...){
    forbid_additional_args(...)
    uncertainly(f, tau, seed = seed, n = n)(h, detectability)
  }
  mutate(grid,
         certain = pmap_dbl(grid, certain),
         uncertain = pmap_dbl(grid, uncertain))
}

results <- mutate(measures, results = map(fun, apply_on_grid, grid)) %>% 
  select(-fun) %>% 
  unnest(results) %>% 
  pivot_longer(c(certain, uncertain),
               names_to = "certainty",
               values_to = "value")
```

```{r, cache = TRUE}
plots <- results %>%
  mutate(name = str_remove(name, "Measure")) %>%
  split(., .$name) %>%
  imap(
    ~ ggplot() +
      geom_line(aes(tau, value, linetype = certainty), data = .x) +
      labs(title = .y,
           y = element_blank(),
           x = "$\\tau$") +
      geom_ribbon(
        data = pivot_wider(.x, names_from = certainty, values_from = value),
        aes(x = tau, ymin = uncertain, ymax = certain),
        alpha = .1
      ) +
      theme_tufte() +
      NULL
  )
```

>>>>

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
output:
  bookdown::pdf_document2:
    latex_engine: "xelatex"
    toc: false
    number_sections: false
header-includes:
   - \usepackage{libertine}
fontsize: 11pt
linestretch: 1.2
geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"
papersize: a4
bibliography: references.bib
abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
csl: apa.csl
repro:
  scripts:
    - R/plots.R
  files:
    - abstract.md
    - data/simulation.rds
  packages:
    - tidyverse
    - ggplot2
    - ggthemes
    - patchwork
    - tikzDevice
  tlmgr:
    - pgf
    - preview
    - libertine
  apt:
    - git
    - rsync
    - curl
    - inkscape
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE
)
knitr::read_chunk("R/plots.R")

@@

@@

```{r}
results <- readr::read_rds(here::here("data", "simulation.rds"))
```

```{r measure-plots}
# see R/plots.R
```
